---
title: "Teradata"
author: "Nikhil Bhatewara"
date: "April 25, 2018"
output: 
  html_document: 
    fig_height: 9
    fig_width: 9
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,out.width='\\textwidth',fig.height=6,fig.align='center')
```

## work space

```{r,opts_chunk$set(out.width='\\textwidth',fig.height=8,fig.align='center')}
### Import libraries
library("plyr")
require("stats")
require("ggplot2")
library("stringdist")
library("dplyr")
library("stringr")
#require("scales")
workingdirectory <- "D:\\e\\Osu\\Spring 2018\\R_python\\teradata\\working\\cleandata"
#workingdirectory="C:\\Users\\msis5223user\\Documents\\teradata_nikhil"
setwd(workingdirectory)
eventsbike_clean <- read.csv("eventsbike_clean.csv")
#summary(eventsbike_clean)
eventsbike_clean <- subset(eventsbike_clean, select = -c(X))




eventsbike_clean[,c(3,4,13,14,27,28,29,21)] <- sapply(eventsbike_clean[,c(3,4,13,14,27,28,29,21)],tolower)


```
- Creating a column "Goal_Achieved" which represents % of goal achieved in each event

```{r}
df0 <- eventsbike_clean[order(-eventsbike_clean$Total_of_All_Confirmed_Gifts),]
df0$Goal_Achieved<-ifelse(df0$Event_Goal==0, 101, (100*(df0$Total_of_All_Confirmed_Gifts/df0$Event_Goal)))
```
Extracting those events which achieved target goal by more than 80%, excluding those events which have goal events < 1 

```{r}
morethan80percent <- (df0$Goal_Achieved > 80 & df0$Event_Goal>1 )
top80events <- df0[morethan80percent,c(2, 7,11,33)]

```
Similarly extracting those events which did not perform well , target goal achieved less than 30%
```{r}
lessthan30percent <- (df0$Goal_Achieved < 30 & df0$Event_Goal>1 )
worst30events <- df0[lessthan30percent,c(2, 7,11,33)]

```
# Analysis on TOP 80% STATES

```{r}
df1 <- df0[df0$Event_ID %in% top80events$Event_ID,c(1,2,27,28,29,30,33)]

```
- TOP 5 STATES BY NUMBER OF EVENTS HAVING > 80 % GOAL ACHIEVED

```{r}
temp1<-plyr::count(df1,"State")
#head(temp1[order(-temp1$freq),],n=5)
ga7 <- (ggplot(temp1,aes(x=reorder(State,-freq),y=freq))+geom_bar(stat='identity',fill='green') + theme_classic())

ga7 + labs(y = "Average Number of Events",x="States",title="Distribution of Number of Events for each state",subtitle="For Top 20% events (Fundraising goal achieved by more than 80%)",caption="Based on data from events file") +guides(fill=FALSE)
```

- TOP 5 STATES BY AVERAGE % GOAL ACHIEVED

```{r}
temp2 <- aggregate(Goal_Achieved~State,data=df1,mean)
#head(temp2[order(-temp2$Goal_Achieved),],n=5)
ga8 <- (ggplot(temp2,aes(x=reorder(State,-Goal_Achieved),y=Goal_Achieved))+geom_bar(stat='identity',fill='green') + theme_classic())

ga8 + labs(y = "Goal achieved (in %)",x="States",title="Distribution of Goal achieved by each state",subtitle="For Top 20% events (Fundraising goal achieved by more than 80%)",caption="Based on data from events file") +guides(fill=FALSE)
```

TX,NY,MN,PA,OH,MI,ID,KS,MA are best performing states and should be retained.

Invest more on less performing states

# Analysis on BOTTOM 30% STATES

```{r}
df2 <- df0[df0$Event_ID %in% worst30events$Event_ID,c(1,2,28,29,30,33)]
```

- TOP 5 STATES BY NUMBER OF EVENTS HAVING < 30 % GOAL ACHIEVED

```{r}
temp3<-plyr::count(df2,"State")
#head(temp3[order(-temp3$freq),],n=5)

ga9 <- (ggplot(temp3,aes(x=reorder(State,-freq),y=freq))+geom_bar(stat='identity',fill='red') + theme_classic())

ga9 + labs(y = "Average Number of Events",x="States",title="Distribution of Number of Events for each state",subtitle="For bottom 30% events based on goal achieved",caption="Based on data from events file") +guides(fill=FALSE)

```

- TOP 5 STATES BY AVERAGE % GOAL ACHIEVED 

```{r}
temp4 <- aggregate(Goal_Achieved~State,data=df2,mean)
#head(temp4[order(-temp4$Goal_Achieved),],n=5)

ga10 <- (ggplot(temp4[1:30,],aes(x=reorder(State,-Goal_Achieved),y=Goal_Achieved))+geom_bar(stat='identity',fill='red') + theme_classic())

ga10 + labs(y = "Goal achieved (in %)",x="States",title="Distribution of Goal achieved by each state",subtitle="For bottom 30% events based on goal achieved",caption="Based on data from events file") +guides(fill=FALSE)
```

NY,PA,CO,MD,OH,IA,NV,WY are not performing well and hence decision decision should be made whether to invest or not

## Notice that NY,PA,OH are part of both list. So these have both successfull events as well as failed events. Reallocate these resources to top 80%


# TOP event ids from top80perc events
```{r}
top80eventsdf <- df0[df0$Event_ID %in% top80events$Event_ID, ]
```
# import teams data
```{r}
teamsbike_clean= read.csv("teamsbike_clean.csv")
#summary(teamsbike_clean)
teamsbike_clean <- subset(teamsbike_clean, select = -c(X))
teamsbike_clean[, c(2, 5, 8, 9, 10, 11, 21, 22)] = sapply(teamsbike_clean[, c(2, 5, 8, 9, 10, 11, 21, 22)], tolower)

```

- Create df3 containing event ids of top80% events

```{r}
df3 <- teamsbike_clean[teamsbike_clean$Event_ID %in% top80events$Event_ID,]
```

- Count no of participants in each fiscal year for top 80% events

```{r}
temp5 <- aggregate(Number_of_Participants~Fiscal_Year,data=df3,FUN=length)
ggplot(temp5,aes(x=Fiscal_Year,y=Number_of_Participants))+geom_bar(stat='identity',fill='blue') + theme_classic()
```
# No of participants for top 80% goal achieving events are decreasing 


- count no of participants of each fiscal year for all the events

```{r}

temp50 <- aggregate(Number_of_Participants~Fiscal_Year,data=teamsbike_clean,FUN=length)
pt <- (ggplot(temp50,aes(x=Fiscal_Year,y=Number_of_Participants))+geom_bar(stat='identity',fill='yellow') + theme_classic()) 
pt + labs(x = "Fiscal Year",title="Distribution of  participants per year",subtitle="For all events",caption="Based on data from teamsbike") +scale_x_continuous(breaks=pretty(teamsbike_clean$Fiscal_Year, n = 6))
```







- Create df4 containing event ids of worst30% events

```{r}
df4 <- teamsbike_clean[teamsbike_clean$Event_ID %in% worst30events$Event_ID,]
```

- Count no of participants in each fiscal year for worst 30% events

```{r}
temp6 <- aggregate(Number_of_Participants~Fiscal_Year,data=df4,FUN=length)
ggplot(temp6,aes(x=Fiscal_Year,y=Number_of_Participants))+geom_bar(stat='identity',fill='cyan') + theme_classic()
```


# Mouth publicity is more effective

```{r}

temp9<-plyr::count(teamsbike_clean,"Team_Division")
(head(temp9[order(-temp9$freq),],n=25))

```

# Analysis on Companies

- Finding companies which have most participation on top80p events and btm30p events. Then extracting those companies which are present in both. Those companies have growth opportunity

```{r}
temp7<-plyr::count(df3,"Company")
lessthan20_top80p <- temp7[temp7$freq < 20,]
temp8<-plyr::count(df4,"Company")
lessthan20_btm30p <- temp8[temp8$freq < 20,]
temp10 <- lessthan20_top80p[lessthan20_top80p$Company %in% lessthan20_btm30p$Company,]
dim(temp10)
```

- Lets remove somellp since it is not useful
```{r}
temp10 <- temp10[!(temp10$Company == 'some llp'),]
unique(temp10$Company)
```
- Now we have companies which has potential to grow, so lets extract their details from original teams dataset

```{r}
temp11 <-   teamsbike_clean$Company %in% temp10$Company   
temp12 <- teamsbike_clean[temp11,]
#colnames(temp12)
```

- Now we want to check if number fof participants in there companies are incereasing or decreasing, so we will aggregate the number of participants w.r.t comapny and fiscal year

```{r}
temp13 <- aggregate(Number_of_Participants~Company+Fiscal_Year,data=temp12,FUN=length)
temp13 <- temp13[order(temp13$Company),]
head(temp13)
```

- Now we will plot the no of participants for each fiscal year for our 18 companies

```{r}



g0 <- (ggplot(
data = temp13,
aes(
x = Fiscal_Year,
y = Number_of_Participants,
group = Company,

color = Company
)
) +
geom_line(size=1) +
scale_linetype_manual(values = seq(1, 19)) +
scale_shape_manual(values = seq(1, 19)))


g0 + labs(y = "Frequency of Events",x="Fiscal Year",title="Number of Events per year",subtitle="For Companies having potential to increase contribution",caption="Based on data from teams file") +guides(fill=TRUE)+ theme_classic()


```

```{r}

g1 <- (ggplot(
data = temp13[1:17, ],
aes(
x = Fiscal_Year,
y = Number_of_Participants,
group = Company,
colour = Company
)
) +
geom_line(aes(linetype = Company), size = 1) +
geom_point(aes(shape = Company, size = 2)) +
scale_linetype_manual(values = seq(1, 19)) +
scale_shape_manual(values = seq(1, 19)))
g1 + labs(y = "Frequency of Events",x="Fiscal Year",title="Number of Events per year",subtitle="For Companies having potential to increase contribution",caption="Based on data from teams file") +guides(fill=TRUE)


g2 <- (ggplot(
data = temp13[18:33, ],
aes(
x = Fiscal_Year,
y = Number_of_Participants,
group = Company,
colour = Company
)
) +
geom_line(aes(linetype = Company), size = 1) +
geom_point(aes(shape = Company, size = 2)) +
scale_linetype_manual(values = seq(1, 19)) +
scale_shape_manual(values = seq(1, 19)))
g2 + labs(y = "Frequency of Events",x="Fiscal Year",title="Number of Events per year",subtitle="For Companies having potential to increase contribution",caption="Based on data from teams file") +guides(fill=TRUE)


g3 <- (ggplot(
data = temp13[34:49, ],
aes(
x = Fiscal_Year,
y = Number_of_Participants,
group = Company,
colour = Company
)
) +
geom_line(aes(linetype = Company), size = 1) +
geom_point(aes(shape = Company, size = 2)) +
scale_linetype_manual(values = seq(1, 19)) +
scale_shape_manual(values = seq(1, 19)))
g3 + labs(y = "Frequency of Events",x="Fiscal Year",title="Number of Events per year",subtitle="For Companies having potential to increase contribution",caption="Based on data from teams file") +guides(fill=TRUE)


g4 <- (ggplot(
data = temp13[50:66, ],
aes(
x = Fiscal_Year,
y = Number_of_Participants,
group = Company,
colour = Company
)
) +
geom_line(aes(linetype = Company), size = 1) +
geom_point(aes(shape = Company, size = 2)) +
scale_linetype_manual(values = seq(1, 19)) +
scale_shape_manual(values = seq(1, 19)))

g4 + labs(y = "Frequency of Events",x="Fiscal Year",title="Number of Events per year",subtitle="For Companies having potential to increase contribution",caption="Based on data from teams file") +guides(fill=TRUE)

```



# What occupations were responsible for most of our fundraising?
```{r}
#import teams data
participantbike_clean = read.csv("participantbike_clean.csv")
#summary(participantbike_clean)
participantbike_clean <- subset(participantbike_clean, select = -c(X))
#colnames(participantbike_clean)
#change everything text to lowercase

participantbike_clean[, c(1, 3, 5, 6, 8, 11, 13, 14, 15, 16, 17, 23, 24, 25, 26, 27, 28,29,31,33,36,37,38,39)] <- sapply(participantbike_clean[, c(1, 3, 5, 6, 8, 11, 13, 14, 15, 16, 17, 23, 24, 25, 26, 27, 28,29,31,33,36,37,38,39)], tolower)

#head(participantbike_clean)

#top80eventsdf contains best events,extraing those event id, then finding the occupation and performing an aggregation 
#will help us answer this question

temp15 <- participantbike_clean$Event_ID %in% top80eventsdf$Event_ID
temp16 <- participantbike_clean[temp15,]
#dim(temp16)

#colnames(temp16)
#extracting only occupations,team_division,total
temp17 <- temp16[,c(18,21,22,25,32,33)]

occupation_count <- plyr::count(temp17,"Participant_Occupation")
occupation_count_df <- as.data.frame(occupation_count)
#Since some occupation could be of different type but belong to same set,use fuzzy matching

df1<-occupation_count$Participant_Occupation
df2<-occupation_count$Participant_Occupation
df1 <- as.data.frame(df1)
df2 <- as.data.frame(df2)
df1$names_from_df2 <- " "
names(df1) <- c("Participant_Occupation","names_from_df2")
#colnames(df1)
names(df2) <- c("Participant_Occupation")
#dim(df1)
#dim(df2)
yournewvector <- lapply(df1$Participant_Occupation,agrep,df2$Participant_Occupation,max.distance=0.2)
#yournewvector

#write.csv(df2, file="df2.csv", row.names=FALSE)

#It is not very useful in our case
```

```{r}
occupation_count <- occupation_count[order(-occupation_count$freq),]
occupation_count_gg <- occupation_count[order(-occupation_count$freq),]
temp19 <- occupation_count_gg[2:7,] #top 5 except unknown
p<- ggplot(data=temp19, aes(x = reorder( Participant_Occupation,freq), y = freq,fill=Participant_Occupation )) + geom_bar(stat ='identity')
p + coord_flip() + labs(y = "Frequency",x="Occupation of participants",title="Most common Occupation of participants",subtitle="For events which acieved more than 80% of fundraising goal",caption="Based on data from participants file") +guides(fill=FALSE)
```
```{r}
occupation_count_forall <- plyr::count(participantbike_clean,"Participant_Occupation")
occupation_count_forall <- occupation_count_forall[order(-occupation_count_forall$freq),]

p<- ggplot(data=occupation_count_forall[2:7,], aes(x = reorder( Participant_Occupation,freq), y = freq,fill=Participant_Occupation )) + geom_bar(stat ='identity')
p + coord_flip() + labs(y = "Count",x="Occupation of participants",title="Most common Occupation of participants",subtitle="For all events",caption="Based on data from participants file") +guides(fill=FALSE)
```



- Engineering,  Healthcare , IT, Sales, Management & Education


# common attributes in top performing corporate teams
```{r}
df5 <- teamsbike_clean[, c(3, 10, 14)]
#unique(df5$Team_Division)
#dim(df5)
#colnames(df5)
temp20 <- df5[df5$Team_Division == 'corporate', ]

#head(temp20)

temp21 <- temp20[order(-temp20$Team_Total_Confirmed), ]

top50corpteams <- head(temp21, 50)

temp22 <-
participantbike_clean[participantbike_clean$Event_ID %in%  top50corpteams$Event_ID, ]

#head(temp22)
```

- by type
```{r}

temp23 <- plyr::count(temp22,"Participation_Type_Name")

temp23 <- temp23[order(-temp23$freq),]

#colnames(temp23)

p1<- ggplot(data=head(temp23[order(-temp23$freq),]), aes(x = reorder( Participation_Type_Name,freq), y = freq,fill=Participation_Type_Name )) + geom_bar(stat ='identity')

p1 + coord_flip()+labs(y = "Frequency",x="Participation Type",title="Most common type of participants",subtitle="For top 50 corporate teams",caption="Based on data from participants & teams file") +guides(fill=FALSE)
```

- by occupation

```{r}

temp24 <- plyr::count(temp22,"Participant_Occupation")
temp24 <- temp24[order(-temp24$freq),]
temp25 <- temp24[!temp24$Participant_Occupation=='unknown',]
p2<- ggplot(data=head(temp25), aes(x = reorder( Participant_Occupation,freq), y = freq,fill=Participant_Occupation )) + geom_bar(stat ='identity')
p2 + coord_flip() + labs(y = "Frequency",x="Occupation of participants",title="Most common Occupation of participants",subtitle="For top 50 corporate teams(excluding unknown occupation)",caption="Based on data from participantsbike file") + guides(fill=FALSE)
```

- by connectiontoms

```{r}
temp26 <- plyr::count(temp22,"Participant_Connection_to_MS")
temp26 <- temp26[order(-temp26$freq),]
p3<- ggplot(data=head(temp26), aes(x = reorder(Participant_Connection_to_MS,freq), y = freq,fill=Participant_Connection_to_MS )) + geom_bar(stat ='identity')
p3 + coord_flip() +labs(y = "Frequency",x="Connection to MS",title="What drive corporate participants towards MS events",subtitle="For top 50 corporate teams",caption="Based on data from participants & teams file") + guides(fill=FALSE)
```
- by State

```{r}
temp27 <- plyr::count(temp22,"Address_._Participant_State.Province")
temp27 <- temp27[order(-temp27$freq),]
p4<- ggplot(data=head(temp27), aes(x = reorder(Address_._Participant_State.Province,freq), y = freq,fill=Address_._Participant_State.Province )) + geom_bar(stat ='identity')
p4 + coord_flip()+labs(y = "Frequency",x="States",title="Where most corporate events occur?",subtitle="For top 50 corporate teams",caption="Based on data from participants & teams file") + guides((fill=FALSE))
```

- by registration type

```{r}
temp28 <- plyr::count(temp22,"Registration_Type")
temp28 <- temp28[order(-temp28$freq),]
#table(temp28)
p5<- ggplot(data=head(temp28), aes(x = reorder(Registration_Type,freq), y = freq,fill=Registration_Type )) + geom_bar(stat ='identity')
p5 + labs(y = "Count",x="Registration Type",title="Which type of registration is preferred by corporate participants",subtitle="For top 50 corporate teams",caption="Based on data from participants file")+guides(fill=FALSE)
```

- by gender

```{r}
temp29 <- plyr::count(temp22,"Participant_Gender")
temp29 <- temp29[order(-temp29$freq),]
#table(temp29)
p6<- ggplot(data=head(temp29), aes(x = reorder(Participant_Gender,freq), y = freq,fill=Participant_Gender )) + geom_bar(stat ='identity')
p6 + labs(y = "Count",x="Gender",title="Distribution of Gender of particiapnts",subtitle="For top 50 corporate teams",caption="Based on data from participant file")+guides(fill=FALSE)
```

- by industry including unknown occupation

```{r}

temp32 <- plyr::count(participantbike_clean, "Industry")
temp32 <- temp32[order(-temp32$freq), ]
p8 <- ggplot(data = head(temp32), aes(x = reorder(Industry, freq), y = freq, fill = Industry)) + geom_bar(stat = "identity")
p8 + coord_flip() +labs(y = "Frequency",x="Industry",title="Which industries attracts most participants",subtitle="For all events (including unknown occupation)",caption="Based on data from participants file")+guides(fill=FALSE)

```

- by industry excluding unknown occupation

```{r}
temp33 <-   filter(participantbike_clean,!Participant_Occupation == "unknown")
#dim(temp33)
temp34 <- plyr::count(temp33, "Industry")
temp34 <- temp34[order(-temp34$freq), ]
p9 <- ggplot(data = head(temp34), aes(x = reorder(Industry, freq), y = freq, fill = Industry)) + geom_bar(stat = "identity")
p9 + coord_flip()  +labs(y = "Frequency",x="Industry",title="Which industries attracts most participants",subtitle="For all events (excluding unknown occupation)",caption="Based on data from participants file")+guides(fill=FALSE)
```


- in events data checking average team size for top corporate teams

```{r}
temp30 <- eventsbike_clean[eventsbike_clean$Event_ID %in% top50corpteams$Event_ID,]
#head(temp30)
temp31 <- plyr::count(temp30,"Average_Team_Size")

p7<- ggplot(data=head(temp31), aes(x = reorder(Average_Team_Size,freq), y = freq,fill=Average_Team_Size )) + geom_bar(stat ='identity')
p7  +labs(y = "Frequency",x="Team size",title="Average team size of participants",subtitle="For top 50 corporate teams",caption="Based on data from events file")+guides(fill=FALSE)
```

# Assessing competitive events

- Bestbuddies Boston Hynnais port
- Bestbuddies San Simone - no presence of MS


```{r}

#colnames(participantbike_clean)
#df for events having in miami
florida <- participantbike_clean[participantbike_clean$Address_._Participant_State.Province=="fl",]
temp100 <- grep("*miami*",florida$Address_._Participant_City)
miami <- florida[temp100,]
#df for events in boston
boston <- participantbike_clean[participantbike_clean$Address_._Participant_City == "boston",]
bestbuddies <- rbind(miami,boston)
mean(bestbuddies$Total_Not_From_Participant)
mean(bestbuddies$Number_Not_From_Participant)

#df for st louis
mo <- participantbike_clean[participantbike_clean$Address_._Participant_State.Province=="mo",]
temp101 <- grep("*st. louis*",mo$Address_._Participant_City)
st_louis <- mo[temp101,]
#df for san diego
sandiego <- participantbike_clean[participantbike_clean$Address_._Participant_City == "san diego",]

pedalthecause <- rbind(st_louis,sandiego)

sum(pedalthecause$Total_of_All_Confirmed_Gifts)




```

MS society vs bestbuddies
raised average $616 by participant (which does not include their own contribution) 

ms vs pealthecause
1808037 donatons received
```{r}
donations <- read.csv("donations2013_2017.csv")

alldonors <- length(unique(donations$Donor_Member_ID))
alldonors

temp500 <- donations[,c('Donor_Member_ID','Fiscal_Year')]
#colnames(temp500)

uniqueindex <- as.numeric(rownames(unique(temp500)[1])) 

temp501 <- temp500[uniqueindex,]

#temp502 <- aggregate(Donor_Member_ID~Fiscal_Year,data=temp501,FUN=length)
temp502 <- aggregate(Donor_Member_ID~Fiscal_Year,data=donations,FUN=length)
p9<- ggplot(data=temp502[2:6,], aes(x = Fiscal_Year, y = Donor_Member_ID,fill= Fiscal_Year )) + geom_bar(stat ='identity')
options(scipen=1000000)
p9  +labs(y = "Number of Unique Donors",x=" Fiscal Year",title="Distribution of Unique Donors",subtitle="For events since 2013",caption="Based on data from donations file")+guides(fill=FALSE)


```
```{r}


#(teamsbike_clean$Number_of_Participants)

temp503 <- aggregate(Number_of_Participants~Fiscal_Year,data=teamsbike_clean,FUN=sum)
p10<- ggplot(data=temp503, aes(x = Fiscal_Year, y = Number_of_Participants,fill= Fiscal_Year )) + geom_bar(stat ='identity')
options(scipen=1000000)
p10  +labs(y = "Number of Participants",x=" Fiscal Year",title="Distribution of Participants",subtitle="For events since 2013",caption="Based on data from teams file")+guides(fill=FALSE)



```

```{r}
temp502 <- temp502[2:6,]
temp503 <- temp503[1:5,]
names(temp502) <- c("Fiscal_Year","Number_of_Donors")

Fiscal_Year<- temp502$Fiscal_Year

donor_participant_ratio <- temp502$Number_of_Donors/temp503$Number_of_Participants



temp504 <- data.frame(Fiscal_Year,donor_participant_ratio)

temp504


p11<- ggplot(data=temp504, aes(x = Fiscal_Year, y = donor_participant_ratio,fill= Fiscal_Year )) + geom_bar(stat ='identity')
options(scipen=1000000)
p11  +labs(y = "Number of donor for each participant",x=" Fiscal Year",title="Donors per participant",subtitle="For events since 2013",caption="Based on data from teams and donations file")+guides(fill=FALSE)


```

Bestbuddies
raised $2553
13 donors

pedalthecause
81000 donars
19809668 total donation

pelotonia
157751378$ raised since 2009
1362303 since in 2018
156389075 2009-2017
$ 19548634 per year

pan mass challenge 2017(http://archive.boston.com/news/local/gallery/080407PanMassChallenge?pg=2)

5000 riders
